How do agents form inflation expectations? evidences from cross-moment estimation and the uncertainty

### 2020-04-10 paper/InfVar.tex
finished the short draft of DE theory.

	This paper extends the abovementioned effort by estimating each theory parametrically accounting for cross-moment restrictions. This is a natural step forward from reduced-form tests of rationality widely seen in the literature. With comparable estimates of different theories, one can evaluate to what extent each theory fits the observed expectation data and inflation dynamics.  I evaluate each model of expectation in terms of its sensitivity of specification with respect to four criteria (1) what moments are used for estimation. (2) the specification of the underlying process of the inflation. (3)  if estimating the underlying process and expectations separately or jointly. The former basicaly assumes recovers the inflation process only based on inflation data while the later let the expectations to provide information used for inflation estimation as well. (4) if it accounts for households and forecasters' expectations equally well.  Broadly speaking, this framework is closer to the original spirit of rational expectation in the sense it focuses on the consistency of the model specification instead of the absolute rationality.

	This is not the first research that estimates theory using cross-moment restrictions (For instance, \cite{giacomini2020heterogeneity}). But one novel aspect of this paper lies in utilizing an additional moment of forecasts in estimation, i.e. the uncertainty. Existing work that studies expectation formation utilizing survey data mostly focuses on the individual and cross-sectional patterns of the mean forecast of the variable. But there are additional insights from the forecast uncertainty, which is only available if each surveyee is asked to assign their own perceived probabilities to a range of values of the variable. For instance, the dispersion of uncertainty across individual forecasters, as that of the average forecasts, are inconsistent with the benchmark full-information rational expectation as the latter assumes that agents agree on the generating process of the data and have the common knowledge of available information. Dynamically, across different vintages of the forecast, the revision in uncertainty is a measure of information gain or the degree of forecasting efficiency. Besides, the relationship between uncertainty with other moments such as forecast errors and disagreements provides a way to check if the surveyed expectations are consistent with certain theories of expectation formation. In addition, although some theories have similar qualitative predictions about the forecasting moments, the configurations of the model parameters consistent with the theory may not be empirically realistic.

	With the framework of estimation presented above, this paper evaluates three workhorse theories on expectation formation. (1) Sticky expectation (SE). (2) Noisy information (NI). (3) Diagnostic expectations (DE).  One major distinction between the first two and DE is that the former predict underreaction due to rigidity and the latter predicts overreaction.  The distinction between SE and NI subtly depends on the quantitative realism of the data implied parameters.

@@

	Different from the previous two theories, diagnostic expectation (\cite{bordalo2018diagnostic}) is developed in the literature to account for extrapolative expectation formation that features overreactions to news. Both SE and NI deviate from FIRE in terms of the information set available to the agents, while DE deviates from FIRE in terms of the process of the otherwise perfect information set.

@@

	For a given process of inflation and a particular theory of expectation formation, the GMM estimate of the vector of parameters $\Omega$ is defined as the following.

@@

	\begin{itemize}
		\item  $\Omega$: parameters of the particular $o \in \{{fire}, {se}, {ni},{de},{seni} \} \times \{ar, sv\}$
		\item  $\Gamma$: constraints for the parameter. 
		\item $M_{data}$: data moments
		\item $F$: simulated model moments according to a particular theory $o$, a function of parameters $\Omega$ as well as the $Y$, the real-time data (including history) up till each point of the time $t$. 
		\begin{itemize}
			\item unconditional moments, not specific to time
			\item moments selected from average forecast, variance and autocovariance of forecasts, average diagreement, variance and autovariance of disagreement, average uncertainty, etc. 
		\end{itemize}
		\item  $W$: weight matrix, identity matrix for now 
	\end{itemize}

>>>>

	This paper extends the abovementioned effort by estimating each theory parametrically accounting for cross-moment restrictions. This is a natural step forward from reduced-form tests of rationality widely seen in the literature. With comparable estimates of different theories, one can evaluate to what extent each theory fits the observed expectation data and inflation dynamics.  I evaluate each model of expectation in terms of its sensitivity of the parameter estimates with respect to four criteria (1) what moments are used for estimation; (2) the specification of the underlying process of the inflation; (3)  if estimating the underlying process and expectations separately or jointly. The former basicaly recovers the inflation process only based on inflation data while the later let the expectations to provide information in estimating the inflation process. (4) If it accounts for households and forecasters' expectations equally well.  Broadly speaking, this framework is closer to the essence of the rational expectation assumption in the sense that it focuses on the consistency of the model specification instead of the absolute rationality.

	This is not the first research that estimates theory using cross-moment restrictions (For instance, \cite{giacomini2020heterogeneity}). But one novel aspect of this paper lies in utilizing an additional moment of forecasts in estimation, i.e. the uncertainty. Existing work that studies expectation formation utilizing survey data mostly focuses on the individual and cross-sectional patterns of the mean forecast and forecast errors. But there are additional insights from the forecast uncertainty, which is only available if each surveyee is asked to assign their own perceived probabilities to a range of values of the variable. For instance, the dispersion of uncertainty across individual forecasters, as that of the average forecasts, are inconsistent with the benchmark full-information rational expectation as the latter assumes that agents agree on the generating process of the data and have the common knowledge of available information. Dynamically, across different vintages of the forecast, the revision in uncertainty is a measure of information gain or the degree of forecasting efficiency. Besides, the relationship between uncertainty with other moments such as forecast errors and disagreements provides a way to check if the surveyed expectations are consistent with certain theories of expectation formation. In addition, although some theories have similar qualitative predictions about the forecasting moments, the configurations of the model parameters consistent with the theory may not be empirically realistic.

	With the framework of estimation presented above, this paper evaluates three workhorse theories on expectation formation. (1) Sticky expectation (SE). (2) Noisy information (NI). (3) Diagnostic expectations (DE).  One major distinction between the first two and DE is that the former predict underreaction due to rigidity and the latter predicts overreaction.  The distinction between SE and NI subtly depends on the quantitative realism of the data-implied parameters.

@@

	Different from the previous two theories, diagnostic expectation (\cite{bordalo2018diagnostic}) is developed in the literature to account for extrapolative expectations that feature overreactions to the news. Both SE and NI deviate from FIRE in terms of the information set available to the agents, while DE deviates from FIRE in terms of the processing of an otherwise fully updated information set.

Omitting the micro foundation, the following equation captures the essence of DE's mechanism. Each individual $i$'s h-period-ahead forecast consists of two components. The first component can be considered as a rational forecast based on the fully updated $y_t$. The second component corresponds to the potential overreaction to the unexpected shock to inflation at time t, i.e. the forecast error. The degree of overreaction is governed by the parameter $\theta$. The premise of DE models is that $\theta \geq 0$. The forecast collapses to the FIRE when $\theta = 0$. Any $\theta>0$ implies cases in which the agent revises her forecast overly to the realized forecast error in the same direction.

 while since agents extrapolates whatever is considered 
		\begin{eqnarray}
		\begin{aligned}
			\bar y_{i,t+h|t}  =  \rho^h y_{t} + \theta_i (\rho^h y_t - \bar y_{i,t+h|t-1})
		\end{aligned}
	\end{eqnarray}

	 I allow $\theta$ to be different across different agents, therefore, I add the subscript $i$ to the parameter. Since agents are equally informed about the realizations of the variable, the only room for disagreement to rise is either difference in the initial information set or the degree of overreaction differs across people. To capture this, I assume $\theta_i$ to follow a normal distribution across the population with a variance of $\sigma_{\epsilon}$. It will be estimated along with $\theta$.

	 Finally as to the uncertainty, since the mechanism of extrapolation in DE does not change the agent's perceived distribution of the future shocks, benchmark DE theory predicts the forecast uncertainty to remain the same as the rational expectation.

@@

	For a given process of inflation and a particular theory of expectation formation, the SMM estimate of the vector of parameters $\Omega$ is defined as the following.

@@

	\begin{itemize}
		\item  $\Omega$: parameters of the particular $o \in \{{fire}, {se}, {ni},{de} \} \times \{ar, sv\}$
		\item  $\Gamma$: constraints for the parameter. 
		\item $M_{data}$: data moments
		\item $F$: simulated model moments according to a particular theory $o$, a function of parameters $\Omega$ as well as the $Y$, the real-time data (including history) up till each point of the time $t$. 
		\begin{itemize}
			\item unconditional moments, not specific to time
			\item moments selected from average forecast, variance and autocovariance of forecasts, average diagreement, variance and autovariance of disagreement, average uncertainty, etc. 
		\end{itemize}
		\item  $W$: weight matrix, identity matrix for now 
	\end{itemize}
